### 2022-07-19 README.Rmd
almost ready

---
output:
  github_document:
    toc: false
    fig_width: 10.08
    fig_height: 6
tags: [r, complexity]
vignette: >
  %\VignetteIndexEntry{README}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
bibliography: bibliography.bib
csl: utils/apa.csl
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# options and parameters
options(digits = 3)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "../../studies/complexity_structure/figures/"
)
```

# The Structure of Chaos: An Empirical Comparison of Fractal Physiology Complexity Indices using NeuroKit2

<!-- # Benchmarking and Analysis of Complexity Measures -->
<!-- # Measuring Chaos: Complexity and Fractal Physiology using NeuroKit2 -->
<!-- # Measuring Chaos with NeuroKit2: An Empirical Comparison of Fractal Physiology Complexity Indices -->

- See the open-access [**preprint**](https://psyarxiv.com/v5tqw/).

<!-- ```{r, child=c('analysis.Rmd')} -->
<!-- ``` -->

## References