### 2022-07-26 manuscript.Rmd

---
title             : "**The Structure of Chaos: An Empirical Comparison of Fractal Physiology Complexity Indices using NeuroKit2**"
shorttitle        : "Structure of Chaos"

author: 
  - name          : "Dominique Makowski"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dmakowski@ntu.edu.sg"
    orcid         : 0000-0001-5375-9967
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Data curation"
      - "Formal Analysis"
      - "Funding acquisition"
      - "Investigation"
      - "Methodology"
      - "Project administration"
      - "Resources"
      - "Software"
      - "Supervision"
      - "Validation"
      - "Visualization"
      - "Writing – original draft"
  - name          : "An Shu Te"
    affiliation   : "1"
    orcid         : 0000-0001-8008-2824
    role:
      - "Software"
      - "Project administration"
      - "Writing – review & editing"
  - name          : "Tam Pham"
    affiliation   : "1"
    orcid         : 0000-0002-6392-6703
    role:
      - "Software"
      - "Writing – review & editing"
  - name          : "Zen J. Lau"
    affiliation   : "1"
    orcid         : 0000-0003-4429-4082
    role:
      - "Software"
      - "Writing – review & editing"
  - name          : "S.H. Annabel Chen"
    corresponding : yes
    affiliation   : "1, 2, 3, 4"
    orcid         : 0000-0002-1540-5516
    role:
      - "Supervision"
      - "Project administration"
      - "Writing – review & editing"

affiliation:
  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "LKC Medicine, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "National Institute of Education, Singapore"
  - id            : "4"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"

authornote: |
  Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore (dom.makowski@gmail.com) and Annabel Chen (AnnabelChen@ntu.edu.sg).
abstract: |
  Complexity quantification, through entropy, information theory and fractal dimension indices, is gaining a renewed traction in psychophsyiology, as new measures with promising qualities emerge from the computational and mathematical advances. Unfortunately, few studies compare the relationship and objective performance of the plethora of existing metrics, in turn hindering reproducibility, replicability, consistency, and clarity in the field. Using the NeuroKit2 Python software, we computed a list of 112 (predominantly used) complexity indices on signals varying in their characteristics (noise, length and frequency spectrum). We then systematically compared the indices by their computational weight, their representativeness of a multidimensional space of latent dimensions, and empirical proximity with other indices. Based on these considerations, we propose that a selection of 12 indices, together representing 85.97% of the total variance of all indices, might offer a parsimonious and complimentary choice in regards to the quantification of the complexity of time series. Our selection includes *CWPEn*, *Line Length (LL)*, *BubbEn*, *MSWPEn*, *MFDFA (Max)*, *Hjorth Complexity*, *SVDEn*, *MFDFA (Width)*, *MFDFA (Mean)*, *MFDFA (Peak)*, *MFDFA (Fluctuation)*, *AttEn*. Elements of consideration for alternative subsets are discussed, and data, analysis scripts and code for the figures are open-source.

keywords          : "chaos, complexity, fractal, physiology, noise"
wordcount         : "3140"

bibliography      : "bibliography.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : yes
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
csl: utils/mdpi.csl

header-includes:
  - \usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
---

```{r, child='analysis.Rmd'}
```

\newpage

## References

::: {#refs custom-style="Bibliography"}
:::